\bf Adaption of the Chumbley Score to matching of bullet striation marks

### 2018-04-21 doc/paper.Rmd
expected failures, additional coarseness

```{r data}
if (!file.exists("../data/all-sigs.rds")) {
  files <- dir("../data/signatures", pattern="csv")
  all <- data.frame()
  for (file in files) {
    tmp <- read.csv(file=file.path(path="../data/signatures", file)) 
    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
    all <- rbind(all, tmp)
  }
  all <- all %>% filter(land1_id < land2_id)
  saveRDS(all, file="../data/all-sigs.rds")
} else {
  all <- readRDS("../data/all-sigs.rds")
}

@@

```{r data-profiles}
if (!file.exists("../data/all-profiles.rds")) {
  files <- dir("../data/profiles/", pattern="csv")
  allp <- data.frame()
  for (file in files) {
    tmp <- read.csv(file=file.path(path="../data/profiles", file)) 
    #print(names(tmp))
    if (length(grep("coarse",names(tmp))) == 0)
      tmp$coarse <- gsub(".*coar-(.*).csv", "\\1", file)
    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
    allp <- rbind(allp, tmp)
  }
  allp <- allp %>% filter(land1_id < land2_id)
  allp <- allp %>% mutate(coarse = as.numeric(gsub("pt", ".", coarse)))
  saveRDS(allp, file="../data/all-profiles.rds")
} else {
  allp <- readRDS("../data/all-profiles.rds")
}

@@

if (!file.exists("../data/invprofiles.rds")) {
  files <- dir("../data/inverse-profiles/", pattern="rds")
  allinv <- data.frame()
  for (file in files) {
    tmp <- readRDS(file=file.path(path="../data/inverse-profiles", file)) 
    allinv <- rbind(allinv, tmp)
  }
  allinv <- allinv %>% filter(land1_id < land2_id)
  allinv <- allinv %>% select(-coarseness)
  saveRDS(allinv, file="../data/invprofiles.rds")
} else {
  allinv <- readRDS("../data/invprofiles.rds")
}

@@

```{r data-sig-coars}
if (!file.exists("../data/all-sig-c.rds")) {
  files <- dir("../data/sig_diff_coarseness/", pattern="csv")
  all_sig_c <- data.frame()
  for (file in files) {
    tmp <- read.csv(file=file.path(path="../data/sig_diff_coarseness/", file)) 
    #print(names(tmp))
    if (length(grep("coarse",names(tmp))) == 0)
      tmp$coarse <- gsub(".*coarse-(.*).csv", "\\1", file)
    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
    all_sig_c <- rbind(all_sig_c, tmp)
  }
  all_sig_c <- all_sig_c %>% filter(land1_id < land2_id)
  all_sig_c <- all_sig_c %>% mutate(coarse = as.numeric(gsub("pt", ".", coarse)))
  saveRDS(all_sig_c, file="../data/all-sig-c.rds")
} else {
  all_sig_c <- readRDS("../data/all-sig-c.rds")
}

@@

# Matching signature
d1<- sigs_graphics.match[which(sigs_graphics.match$land_id==1,arr.ind = TRUE),]
d2<- sigs_graphics.match[which(sigs_graphics.match$land_id==276,arr.ind = TRUE),]
data1<-d1 %>%select(l30)
#data1<- data1$y
data2<- d2 %>%select(l30)
#data2<- data2$y
window_opt = 120 
window_val = 50
coarse = 1
data1<- matrix(unlist(data1))
data2<- matrix(unlist(data2))

@@

sigs1.min = sigs_graphics[[1]] %>% filter(land_id==1) %>% select(y) %>% min()
sigs2.min = sigs_graphics[[1]] %>% filter(land_id==276) %>% select(y) %>% min()
window_val <- 3*window_val*0.645

@@

# Matching signature
#gi<-
#sigs_graphics.match<- sigs_graphics.match %>% filter(land_id == c(1,9))  
#sigs_graphics.match<- sigs_graphics.match %>% select(y, l30, land_id)
d1<- sigs_graphics.match[which(sigs_graphics.match$land_id==1,arr.ind = TRUE),]
d2<- sigs_graphics.match[which(sigs_graphics.match$land_id==19,arr.ind = TRUE),]
data1<-d1 %>%select(l30)
#data1<- data1$y
data2<- d2 %>%select(l30)
#data2<- data2$y
window_opt = 120 
window_val = 50
coarse = 1
data1<- matrix(unlist(data1))
data2<- matrix(unlist(data2))

@@

The number of same-shift pairs will be zero, if the optimal locations $t_1^{o}$ and $t_2^{o}$ are so far apart, that no segments of size $w_v$ are left on the same sides of the optimal locations, i.e. $t_1^{o} < w_v$ and $t_2^{o} > T_2-w_o-w_v$ or  $t_1^{o} < T_1- w_o - w_v$ and $t_2^{o} < w_v$. While we can assume that once (sub-)class characteristics are removed, optimal locations $t_i^{o}$ are uniformly distributed across the length of the profile, we cannot assume that $t_1^o$ and $t_2^o$ are independent of each other. In particular, for same-source profiles, we would expect a strong dependency between these locations. However, for different source matches, an argument of independence in the locations could be made. In that case, we expect a test to fail with a probability of $\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}$. For an average length of $T_i$ of 1200 pixels, $w_o = 120$ pixels and $w_v = 30$ pixels this probability is about 0.0015.

@@

Initially, the default settings suggested in \citet{hadler} were used: $w_o = 120$ pixels or about 190 $\mu m$ (ten percent of the average length of profiles) and coarseness $c = 0.25$. 
\autoref{fig:fails} shows the percentage of failed tests among the 85,491 land-to-land comparisons of the NIST data for different values of the validation window size $w_v$. For same-source lands up to 12.5 percent of the tests fail using CS1. The highest percentage of failed tests under CS2 is 1.3% for different source tests using a validation window size $w_v$ of 60 pixels.

```{r fails, echo=FALSE, fig.height = 3, fig.width = 7, out.width='0.7\\textwidth', fig.cap="Percent of failed land-to-land comparisons using an optimization window $w_o = 120$ and a coarseness of $c = 0.25$. With an increase in the size of the validation window  a higher percentage of tests fails under both methods (CS1) and (CS2), but the percentage of failed tests is much smaller under (CS2)."}
allinv$method <- "CS2"
allp$method <- "CS1"

prof_fails <- rbind(
  allinv %>% select(wo, wv, method, p_value, match),
  allp %>% filter(coarse == 0.25) %>% 
    select(wo, wv, method, p_value, match)
)
fails <- prof_fails %>%
  group_by(wo, wv, method, match) %>% summarize(
  failed = sum(is.na(p_value)),
  n=n()
)

fails %>% 
  ggplot(aes(x = wv, y = failed/n*100, colour=factor(match), shape=factor(match))) +
  geom_point(size=3) +
  xlab("Size of validation window in pixels") +
  ylab("Percent of failed tests") +
  theme_bw() +
  scale_colour_brewer("Same-source", palette="Set1") +
  scale_shape_discrete("Same-source") +
  ylim(c(0,13)) +
  facet_wrap(~method, labeller="label_both")
```

\hh{Coarseness is another parameter that has to be considered in the matching of bullet land profiles.}

\hh{\autoref{fig:profile-sketch} gives an overview of the effect of different coarseness parameters: from left to right, coarseness levels $c$ are 0.2, 0.25, and 0.3, respectively. The top row shows the resulting signature after smoothing the profile shown in Figure~\ref{fig:sigs-profiles} with the coarseness specified. The histograms in the bottom row show the relative optimal location $t^o$ (optimal location $t^o$ divided by the length of the corresponding profile). Optimal locations should be distributed uniformly. However, for coarseness values of $c = 0.25$ and $c=0.30$ we see very distinct boundary effects: optimal locations $t^o$ are found at the very extreme ends of a profile more often than one would expect based on a uniform location.}

@@

profc20 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30-c_20.rds")
profc20$coarseness <- 0.2
profc25 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30-c_25.rds")
profc30 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30.rds")
profc30$coarseness <- .30
profc <- bind_rows(profc20, profc25, profc30)
profc <- profc %>% mutate(
  locations = chumbley %>% purrr::map_int(.f = function(x) {
    res <- NULL
    try(res <- x$locations[1], silent = TRUE)
    res
  })
)
profsumm <- read.csv("../data/profiles-summary.csv")
profsumm <- profsumm %>% select(-run_id)
profc <- profc %>% left_join(profsumm, by=c("land2_id"="land_id", "profile2_id"="profile_id"))

p2 <- profc %>% 
  ggplot(aes(x = locations/(length-wo))) +
  geom_histogram(binwidth = 0.01) +
  facet_grid(.~coarseness, labeller="label_both") + 
  theme_bw() +
  xlab(expression("Relative optimal location "~t^"*")) +
  ylab("Number of profiles")

@@

\hh{Figure \ref{fig:coarse} shows the type II error rates for profiles using an optimization window $w_o= 120$ and a validation window $w_v = 30$ for varying level of coarseness. The type II error for all nominal levels of  $\alpha$ are the lowest for a coarseness range of 0.20 to 0.30, with a minimum in type II reached at a coarseness $c = 0.25$. }
\hh{However, for an optimal value of coarseness other aspects have to be condsidered.}
```{r coarse, fig.cap="Type II error with respect to coarseness parameter over profiles, $w_o = 120$, $w_v = 30$. Optimal values for coarseness are around $c \\approx 0.30$", fig.width=5, fig.height = 3, out.width='.6\\textwidth'}
e1<-errors2 %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
  geom_point() + geom_line() +
  theme_bw() +
  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Profile Coarseness")+
  ylab("Type II error") 
e1
```

@@

dt <- rbind(wo120_sig, wo120_pr) %>% 
  ungroup() %>% 
  mutate(beta = round(100*beta, 1)) %>%
  select(wv, type, alpha, beta) %>% 
  spread(alpha, beta)
print(xtable(dt), include.rownames=FALSE)
```

>>>>

```{r data}
if (!file.exists("../data/all-sigs.rds")) {
  files <- dir("../data/signatures", pattern="csv")
  all <- data.frame()
  for (file in files) {
    tmp <- read.csv(file=file.path(path="../data/signatures", file)) 
    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% dplyr::select(-x)
    all <- rbind(all, tmp)
  }
  all <- all %>% filter(land1_id < land2_id)
  saveRDS(all, file="../data/all-sigs.rds")
} else {
  all <- readRDS("../data/all-sigs.rds")
}

@@

```{r data-profiles}
if (!file.exists("../data/all-profiles.rds")) {
  files <- dir("../data/profiles/", pattern="csv")
  allp <- data.frame()
  for (file in files) {
    tmp <- read.csv(file=file.path(path="../data/profiles", file)) 
    #print(names(tmp))
    if (length(grep("coarse",names(tmp))) == 0)
      tmp$coarse <- gsub(".*coar-(.*).csv", "\\1", file)
    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% dplyr::select(-x)
    allp <- rbind(allp, tmp)
  }
  allp <- allp %>% filter(land1_id < land2_id)
  allp <- allp %>% mutate(coarse = as.numeric(gsub("pt", ".", coarse)))
  saveRDS(allp, file="../data/all-profiles.rds")
} else {
  allp <- readRDS("../data/all-profiles.rds")
}

@@

if (!file.exists("../data/invprofiles.rds")) {
  files <- dir("../data/inverse-profiles/", pattern="rds")
  allinv <- data.frame()
  for (file in files) {
    tmp <- readRDS(file=file.path(path="../data/inverse-profiles", file)) 
    allinv <- rbind(allinv, tmp)
  }
  allinv <- allinv %>% filter(land1_id < land2_id)
  allinv <- allinv %>% dplyr::select(-coarseness)
  saveRDS(allinv, file="../data/invprofiles.rds")
} else {
  allinv <- readRDS("../data/invprofiles.rds")
}

@@

```{r data-sig-coars}
if (!file.exists("../data/all-sig-c.rds")) {
  files <- dir("../data/sig_diff_coarseness/", pattern="csv")
  all_sig_c <- data.frame()
  for (file in files) {
    tmp <- read.csv(file=file.path(path="../data/sig_diff_coarseness/", file)) 
    #print(names(tmp))
    if (length(grep("coarse",names(tmp))) == 0)
      tmp$coarse <- gsub(".*coarse-(.*).csv", "\\1", file)
    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% dplyr::select(-x)
    all_sig_c <- rbind(all_sig_c, tmp)
  }
  all_sig_c <- all_sig_c %>% filter(land1_id < land2_id)
  all_sig_c <- all_sig_c %>% mutate(coarse = as.numeric(gsub("pt", ".", coarse)))
  saveRDS(all_sig_c, file="../data/all-sig-c.rds")
} else {
  all_sig_c <- readRDS("../data/all-sig-c.rds")
}

@@

# Matching signature
d1<- sigs_graphics.match[which(sigs_graphics.match$land_id==1,arr.ind = TRUE),]
d2<- sigs_graphics.match[which(sigs_graphics.match$land_id==276,arr.ind = TRUE),]
data1<-d1 %>%dplyr::select(l30)
#data1<- data1$y
data2<- d2 %>%dplyr::select(l30)
#data2<- data2$y
window_opt = 120 
window_val = 50
coarse = 1
data1<- matrix(unlist(data1))
data2<- matrix(unlist(data2))

@@

sigs1.min = sigs_graphics[[1]] %>% filter(land_id==1) %>%
  dplyr::select(y) %>% min()
sigs2.min = sigs_graphics[[1]] %>% filter(land_id==276) %>% 
  dplyr::select(y) %>% min()
window_val <- 3*window_val*0.645

@@

# Matching signature
#gi<-
#sigs_graphics.match<- sigs_graphics.match %>% filter(land_id == c(1,9))  
#sigs_graphics.match<- sigs_graphics.match %>% select(y, l30, land_id)
d1<- sigs_graphics.match[which(sigs_graphics.match$land_id==1,arr.ind = TRUE),]
d2<- sigs_graphics.match[which(sigs_graphics.match$land_id==19,arr.ind = TRUE),]
data1<-d1 %>%dplyr::select(l30)
#data1<- data1$y
data2<- d2 %>%dplyr::select(l30)
#data2<- data2$y
window_opt = 120 
window_val = 50
coarse = 1
data1<- matrix(unlist(data1))
data2<- matrix(unlist(data2))

@@

The number of same-shift pairs will be zero, if the optimal locations $t_1^{o}$ and $t_2^{o}$ are so far apart, that no segments of size $w_v$ are left on the same sides of the optimal locations, i.e. $t_1^{o} < w_v$ and $t_2^{o} > T_2-w_o-w_v$ or  $t_1^{o} < T_1- w_o - w_v$ and $t_2^{o} < w_v$, i.e. we have a failure rate of 
\[
P\left( t_1^{o} < w_v \ \cap \ t_2^{o} > T_2-w_o-w_v\right) + P\left( t_1^{o} < T_1- w_o - w_v \ \cap \ t_2^{o} < w_v\right).
\]
While we can assume that once (sub-)class characteristics are removed, optimal locations $t_i^{o}$ are uniformly distributed across the length of the profile, we cannot assume that $t_1^o$ and $t_2^o$ are independent of each other. In particular, for same-source profiles, we would expect a strong dependency between these locations, in which case a large difference between locations is unlikely. However, for different source matches, we can assume that  locations are independent.
In that case, we expect a test to fail with a probability of $\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}$. For an average length of $T_i$ of 1200 pixels, $w_o = 120$ pixels and $w_v = 30$ pixels this probability is about 0.0015.

@@

Initially, the default settings suggested in \citet{hadler} were used: $w_o = 120$ pixels or about 190 $\mu m$ (ten percent of the average length of profiles) and coarseness $c = 0.25$. 
\autoref{fig:fails} shows the percentage of failed tests among the 85,491 land-to-land comparisons of the NIST data for different values of the validation window size $w_v$. For same-source lands up to 12.5 percent of the tests fail using CS1. The highest percentage of failed tests under CS2 is 1.3% for different source tests using a validation window size $w_v$ of 60 pixels.  \hh{Rates of expected failures are based on simulation runs using covariances between locations of same-source profiles of 0.854, and 0.120 for locations from different-source profiles, matching observed covariances for the Hamby scans. }\hh{ Observed failure rates are higher than expected. This might be due to remaining sub-class structure at a coarseness of 0.3 resulting in a distribution of optimal locations different from the assumed uniform. }

```{r simulation, echo=FALSE}
library(MASS) 
Sigsame <- matrix(c(1, 0.85, 0.85, 1), nrow = 2) 
Sigdiff <- matrix(c(1, 0.12, 0.12, 1), nrow = 2) 
set.seed(20140501)
simu <- 1:10 %>% purrr::map_df(.f = function(i) {
  X1 <- mvrnorm(10000, mu = c(0,0), Sigma = Sigsame) 
  X2 <- mvrnorm(10000, mu = c(0,0), Sigma = Sigdiff) 
  Y1 <- pnorm(X1) 
  Y2 <- pnorm(X2) 
  wv <- 10*1:6

  ps <- wv %>% purrr::map_df(.f = function(x) {
    p <- x/(1200-120)
    y1 <- Y1[,1]
    y2 <- Y1[,2]
    difffail1 <- sum((y1 < p & y2 < p) | (y1 > 1-p & y2 > 1-p))/length(y1)
    samefail1 <- sum((y1 < p & y2 >1- p) | (y1 > 1-p & y2 < p))/length(y1)
    y1 <- Y2[,1]
    y2 <- Y2[,2]
    difffail2 <- sum((y1 < p & y2 < p) | (y1 > 1-p & y2 > 1-p))/length(y1)
    samefail2 <- sum((y1 < p & y2 > 1-p) | (y1 > 1-p & y2 < p))/length(y1)
    data.frame(wv=x, samefail1=samefail1, samefail2=samefail2,
               difffail1=difffail1, difffail2=difffail2)
  })
  ps$i <- i
  ps
})
simu$`TRUE` <- simu$samefail1+simu$difffail1
simu$`FALSE` <- simu$samefail2+simu$difffail2
simulong <- simu %>% dplyr::select(wv, i, `TRUE`, `FALSE`) %>% 
  gather(match, failed, 3:4)
simusumm <- simulong %>% group_by(wv, match) %>%
  summarize(
    sdfailed = sd(failed),
    failed = mean(failed)
  )
simusumm$method = "CS1"
```

```{r fails, echo=FALSE, fig.height = 3, fig.width = 7, out.width='0.7\\textwidth', fig.cap="Percent of failed land-to-land comparisons using an optimization window $w_o = 120$ and a coarseness of $c = 0.25$. With an increase in the size of the validation window  a higher percentage of tests fails under both methods (CS1) and (CS2), but the percentage of failed tests is much smaller under (CS2). Observed failure rates are higher than expected rates."}
allinv$method <- "CS2"
allp$method <- "CS1"

allinv <-  allinv %>% mutate(
  loc1 = chumbley %>% purrr::map_int(.f = function(x) x$locations[1]),
  loc2 = chumbley %>% purrr::map_int(.f = function(x) x$locations[2])
)

prof_fails <- rbind(
  allinv %>% dplyr::select(wo, wv, method, p_value, match),
  allp %>% filter(coarse == 0.25) %>% 
    dplyr::select(wo, wv, method, p_value, match)
)
fails <- prof_fails %>%
  group_by(wo, wv, method, match) %>% summarize(
  failed = sum(is.na(p_value)),
  n=n()
)

fails %>% 
  ggplot(aes(x = wv, y = failed/n*100, colour=factor(match), shape=factor(match))) +
  geom_point(aes(size="Observed")) +
  xlab("Size of validation window in pixels") +
  ylab("Percent of failed tests") +
  theme_bw() +
  scale_colour_brewer("Same-source", palette="Set1") +
  scale_shape_discrete("Same-source") +
  ylim(c(0,13)) +
  facet_wrap(~method, labeller="label_both") +
  geom_point(aes(y = failed*100, size = "Expected"), data=simusumm) +
  scale_size_manual("Failures", values=c(1.5,3))
```

## Coarseness
The purpose of the coarseness parameter is to remove (sub-)class characteristics from profiles before comparisons for matching.
\citet{hadler} suggest a coarseness parameter of 0.25 in the setting of toolmark comparisons. For bullet lands, coarseness might need to be adjusted because of the strong effect  bullet curvature has on profiles.

\hh{\autoref{fig:profile-sketch} gives an overview of the effect of different coarseness parameters: from left to right, coarseness levels $c$ are varied in steps of 0.05 from 0.1 to 0.3. The top row shows  resulting signature after smoothing the profile shown in Figure~\ref{fig:sigs-profiles} with the coarseness specified. 
The histograms in the bottom row show the relative optimal location $t^*$. Optimal locations are distributed uniformly once (sub-)class characteristics are removed. However, for coarseness values of $c > 0.20$ we see quite distinct boundary effects: optimal locations $t^*$ are found at the very extreme ends of a profile more often than one would expect based on a uniform distribution.}

@@

profc10 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30-c_10.rds")
profc15 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30-c_15.rds")
profc20 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30-c_20.rds")
profc20$coarseness <- 0.2
profc25 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30-c_25.rds")
profc30 <- readRDS("../data/profiles-coarseness/chumbley-csafe-profiles-inv-wo_160-wv_30.rds")
profc30$coarseness <- .30
profc <- bind_rows(profc10, profc15, profc20, profc25, profc30)
profc <- profc %>% mutate(
  locations = chumbley %>% purrr::map_int(.f = function(x) {
    res <- NULL
    try(res <- x$locations[1], silent = TRUE)
    res
  })
)
profsumm <- read.csv("../data/profiles-summary.csv")
profsumm <- profsumm %>% dplyr::select(-run_id)
profc <- profc %>% left_join(profsumm, by=c("land2_id"="land_id", "profile2_id"="profile_id"))

p2 <- profc %>% 
  ggplot(aes(x = locations/(length-wo))) +
  geom_histogram(binwidth = 0.025) +
  facet_grid(.~coarseness, labeller="label_both") + 
  theme_bw() +
  xlab(expression("Relative optimal location "~t^"*")) +
  ylab("Number of profiles")

@@

\hh{Figure \ref{fig:coarse} shows the type II error rates for profiles using an optimization window $w_o= 120$ and a validation window $w_v = 30$ for varying level of coarseness. The type II error for all nominal levels of  $\alpha$ are the lowest for a coarseness range of 0.20 to 0.30, with a minimum in type II reached at a coarseness $c = 0.25$. The histograms suggest to use a value of coarseness below 0.2.}
\hh{However, for an optimal value of coarseness other aspects have to be condsidered.}
```{r coarse, fig.cap="Type II error with respect to coarseness parameter over profiles, $w_o = 120$, $w_v = 30$. Optimal values for coarseness are around $c \\approx 0.30$", fig.width=5, fig.height = 3, out.width='.6\\textwidth'}
e1<-errors2 %>% filter(wv==30)%>% ggplot(aes(x = coarse, y=beta, colour=factor(alpha))) +
  geom_point() + geom_line() +
  theme_bw() +
  scale_colour_brewer(expression("Nominal\ntype I error"~alpha), palette="Set2") + xlab("Profile Coarseness")+
  ylab("Type II error") 
e1
```

@@

dt <- rbind(wo120_sig, wo120_pr) %>% 
  ungroup() %>% 
  mutate(beta = round(100*beta, 1)) %>%
  dplyr::select(wv, type, alpha, beta) %>% 
  spread(alpha, beta)
print(xtable(dt), include.rownames=FALSE)
```

### 2018-04-21 doc/paper.tex
expected failures, additional coarseness

The number of same-shift pairs will be zero, if the optimal locations
\(t_1^{o}\) and \(t_2^{o}\) are so far apart, that no segments of size
\(w_v\) are left on the same sides of the optimal locations, i.e.
\(t_1^{o} < w_v\) and \(t_2^{o} > T_2-w_o-w_v\) or
\(t_1^{o} < T_1- w_o - w_v\) and \(t_2^{o} < w_v\). While we can assume
that once (sub-)class characteristics are removed, optimal locations
\(t_i^{o}\) are uniformly distributed across the length of the profile,
we cannot assume that \(t_1^o\) and \(t_2^o\) are independent of each
other. In particular, for same-source profiles, we would expect a strong
dependency between these locations. However, for different source
matches, an argument of independence in the locations could be made. In
that case, we expect a test to fail with a probability of
\(\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}\). For an average length of \(T_i\)
of 1200 pixels, \(w_o = 120\) pixels and \(w_v = 30\) pixels this
probability is about 0.0015.

@@

Initially, the default settings suggested in \citet{hadler} were used:
\(w_o = 120\) pixels or about 190 \(\mu m\) (ten percent of the average
length of profiles) and coarseness \(c = 0.25\). \autoref{fig:fails}
shows the percentage of failed tests among the 85,491 land-to-land
comparisons of the NIST data for different values of the validation
window size \(w_v\). For same-source lands up to 12.5 percent of the
tests fail using CS1.
\hh{XXX Currently running CS2 for these settings for a comparison.}

@@

\caption{Percent of failed land-to-land comparisons using an optimization window $w_o = 120$ and a coarseness of $c = 0.25$. With an increase in the size of the validation window  a higher percentage of tests fails. The percentage of failures is higher for same-source lands than for different source lands.}\label{fig:fails}
\end{figure}

\hh{XXX Conclusion: the number of failed tests goes down - a lot. XXX}

\hh{Coarseness is another parameter that has to be considered in the matching of bullet land profiles.}

\hh{\autoref{fig:profile-sketch} gives an overview of the effect of different coarseness parameters: from left to right, coarseness levels $c$ are 0.2, 0.25, and 0.3, respectively. The top row shows the resulting signature after smoothing the profile shown in Figure~\ref{fig:sigs-profiles} with the coarseness specified. The histograms in the bottom row show the relative optimal location $t^o$ (optimal location $t^o$ divided by the length of the corresponding profile). Optimal locations should be distributed uniformly. However, for coarseness values of $c = 0.25$ and $c=0.30$ we see very distinct boundary effects: optimal locations $t^o$ are found at the very extreme ends of a profile more often than one would expect based on a uniform location.}

@@

\hh{Figure \ref{fig:coarse} shows the type II error rates for profiles using an optimization window $w_o= 120$ and a validation window $w_v = 30$ for varying level of coarseness. The type II error for all nominal levels of  $\alpha$ are the lowest for a coarseness range of 0.20 to 0.30, with a minimum in type II reached at a coarseness $c = 0.25$. }
\hh{However, for an optimal value of coarseness other aspects have to be condsidered.}

>>>>

The number of same-shift pairs will be zero, if the optimal locations
\(t_1^{o}\) and \(t_2^{o}\) are so far apart, that no segments of size
\(w_v\) are left on the same sides of the optimal locations, i.e.
\(t_1^{o} < w_v\) and \(t_2^{o} > T_2-w_o-w_v\) or
\(t_1^{o} < T_1- w_o - w_v\) and \(t_2^{o} < w_v\), i.e.~we have a
failure rate of \[
P\left( t_1^{o} < w_v \ \cap \ t_2^{o} > T_2-w_o-w_v\right) + P\left( t_1^{o} < T_1- w_o - w_v \ \cap \ t_2^{o} < w_v\right).
\] While we can assume that once (sub-)class characteristics are
removed, optimal locations \(t_i^{o}\) are uniformly distributed across
the length of the profile, we cannot assume that \(t_1^o\) and \(t_2^o\)
are independent of each other. In particular, for same-source profiles,
we would expect a strong dependency between these locations, in which
case a large difference between locations is unlikely. However, for
different source matches, we can assume that locations are independent.
In that case, we expect a test to fail with a probability of
\(\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}\). For an average length of \(T_i\)
of 1200 pixels, \(w_o = 120\) pixels and \(w_v = 30\) pixels this
probability is about 0.0015.

@@

Initially, the default settings suggested in \citet{hadler} were used:
\(w_o = 120\) pixels or about 190 \(\mu m\) (ten percent of the average
length of profiles) and coarseness \(c = 0.25\). \autoref{fig:fails}
shows the percentage of failed tests among the 85,491 land-to-land
comparisons of the NIST data for different values of the validation
window size \(w_v\). For same-source lands up to 12.5 percent of the
tests fail using CS1. The highest percentage of failed tests under CS2
is 1.3\% for different source tests using a validation window size
\(w_v\) of 60 pixels.
\hh{Rates of expected failures are based on simulation runs using covariances between locations of same-source profiles of 0.854, and 0.120 for locations from different-source profiles, matching observed covariances for the Hamby scans. }\hh{ Observed failure rates are higher than expected. This might be due to remaining sub-class structure at a coarseness of 0.3 resulting in a distribution of optimal locations different from the assumed uniform. }

@@

\caption{Percent of failed land-to-land comparisons using an optimization window $w_o = 120$ and a coarseness of $c = 0.25$. With an increase in the size of the validation window  a higher percentage of tests fails under both methods (CS1) and (CS2), but the percentage of failed tests is much smaller under (CS2). Observed failure rates are higher than expected rates.}\label{fig:fails}
\end{figure}

\subsection{Coarseness}\label{coarseness}

The purpose of the coarseness parameter is to remove (sub-)class
characteristics from profiles before comparisons for matching.
\citet{hadler} suggest a coarseness parameter of 0.25 in the setting of
toolmark comparisons. For bullet lands, coarseness might need to be
adjusted because of the strong effect bullet curvature has on profiles.

\hh{\autoref{fig:profile-sketch} gives an overview of the effect of different coarseness parameters: from left to right, coarseness levels $c$ are varied in steps of 0.05 from 0.1 to 0.3. The top row shows  resulting signature after smoothing the profile shown in Figure~\ref{fig:sigs-profiles} with the coarseness specified. 
The histograms in the bottom row show the relative optimal location $t^*$. Optimal locations are distributed uniformly once (sub-)class characteristics are removed. However, for coarseness values of $c > 0.20$ we see quite distinct boundary effects: optimal locations $t^*$ are found at the very extreme ends of a profile more often than one would expect based on a uniform distribution.}

@@

\hh{Figure \ref{fig:coarse} shows the type II error rates for profiles using an optimization window $w_o= 120$ and a validation window $w_v = 30$ for varying level of coarseness. The type II error for all nominal levels of  $\alpha$ are the lowest for a coarseness range of 0.20 to 0.30, with a minimum in type II reached at a coarseness $c = 0.25$. The histograms suggest to use a value of coarseness below 0.2.}
\hh{However, for an optimal value of coarseness other aspects have to be condsidered.}
